Maximizing Ink in Symmetric Partial Edge Drawings of $k$-plane Graphs

### 2018-01-09 eurocg/paper.tex
zwischenstand

\martin{$f_s$ mit $\ell_i$ linken}

@@

\begin{proof}
	We reduce from the \NP-hard problem \ppsat~\cite{l-pftu-82} using similar ideas as in Bruckdorfer's general proof of the hardness of \maxsped~\cite{b-sgh-15}. %, which reduces from the more restricted problem \textsc{Planar Positive 1-in-3 Sat}. 
	Let $\phi$ be a planar 3-Sat formula with $n$ variables $\{x_1, \dots, x_n\}$ and $m$ clauses $\{c_1, \dots, c_m\}$, each consisting of three literals.
	We can assume that $\phi$ comes with a planar drawing of its variable-clause graph $H_\phi$ that has a vertex for each variable $x_i$ placed on a horizontal line and a vertex for each clause $c_j$ either above or below the horizontal line and connected on that side of the line to the three variables appearing in $c_j$. 
	In our gadget-based reduction we mimic the drawing of $H_\phi$ by creating a 3-plane drawing $\Gamma_\phi$ as a set of line segments of uniform length and a value $L$ such that $\Gamma_\phi$ has a SPED with ink at least $L$ if and only if $\phi$ is satisfiable.
	Figure~\ref{fig:gadget_connection} shows a sketch of the gadgets.

	\begin{figure}
		\centering
		\includegraphics{Reduction_visual}
		\caption{Example of three variable gadgets and a satisfied clause gadget. Dotted parts do not belong to the SPED.}
		\label{fig:gadget_connection}
	\end{figure}

	All segments in the gadgets are of length 5.
	We use pairs of intersecting segments, alternatingly colored red and green.
	The intersection point of each red-green segment pair is at distance 1 from an endpoint.
	Thus the maximum amount of ink contributed by such a pair is 7 (one full segment of length 5 and the other one with two stubs of length 1 each).

@@

	Conversely, assume that we have a SPED with ink value $L$. 
	By construction, every red-green segment pair and every clause gadget must contribute its respective maximum ink value.
	In particular, each variable gadget is either in state \emph{true} or \emph{false}.
	We claim that this truth assignment satisfies $\phi$.
	For each clause, one segment must have full length and as argued before the path connecting that segment to a variable gadget must have all segments at even positions as length-1 stubs and all segments at odd positions as full segments. 
	In particular the first segment of the literal path is fully drawn and therefore in a SPED with ink $L$ the intersecting segment of the variable cannot be fully drawn.
	In fact, it must be drawn as length-1 stubs for otherwise the variable gadget would not contribute its maximum ink value.
	But then by construction the selected variable value sets our literal to \emph{true}.
\end{proof}

@@

Section~\ref{sec:hardness} showed that \maxsped is generally \NP-hard for $k \ge 3$. In this section we consider the special case that the intersection graph of the $k$-plane input drawing is a tree or a cactus. In both cases we present polynomial-time dynamic programming algorithms. Let $ C = (V,E) $ be the intersection graph of a given drawing $\Gamma$ of graph $ G $ as defined in Section~\ref{sec:preliminaries}. Let $ u \in V $ and $ \deg(u) = \delta $. 
Then for the corresponding segment $ s(u) \in S $ there are $ \delta + 1 $ possible stub pairs including drawing the whole segment, see Fig.~\ref{fig:tree}. We store the lengths of the possible stub pairs as node weights $ \ell_1(u), \dots, \ell_\delta(u) \in \mathbb{R}_+$ and assume that $ \ell_1(u),\dots,\ell_\delta(u) $ are sorted from shorter to longer stubs. We define  $ \ell_0(u) $ as the length of the whole segment $ s(u) $.

\begin{figure}[tbp]
	\centering
 	\includegraphics{tree_illustration}
	\caption{A segment $s(u)$ with five intersecting segments and the different induced stub lengths. The boxed stub lengths are considered in $\solshort(u)$.} % The shaded areas are the stubs, $ p(u) $ is the parent of $ u $. The stubs in the boxed area contribute to $ \solshort(u) $.}
	\label{fig:tree}
\end{figure}

@@

Figure~\ref{fig:tree} highlights the stub lengths that contribute to $\solshort(u)$. We recursively define % for a vertex $ u $ and stub length $ \ell_i(u) $:

@@

Recurrence~(\ref{rec:tree}) can be solved naively in $ O(mk^2) $ time. 
Using the order on the stub lengths  we can improve this to $ O(mk) $ time by computing all $ T_i(u) $ for one $ u \in V $ in $O(k)$ time.
Let $u \in V$ be a vertex with degree $\deg(u) = \delta$. 
The values $T_0(u) = \ell_0(u) + \sum_{v\in c(u)} \solshort(v)$ and $T_1(u) = \ell_1(u) + \sum_{v\in c(u)} \sollong(v)$ for the whole segment $s(u)$ and the shortest stubs can be computed in $O(k)$ time each.
Now $T_{j+1}(u)$ can be computed from $T_j(u)$ in $O(1)$ time as follows.
Let $v_j $ be the neighbor of $u$ that induces stub length $\ell_j(u)$ and assume $v_j \ne p(u)$.
In $T_j(u)$ we could still count the value $\sollong(v_j)$, but in $T_{j+1}(u)$ the stub length of $u$ implies that $v_j$ can contribute only $\solshort(v_j)$. 
Then $T_{j+1}(u) = T_j(u) - \sollong(v_j) + \solshort(v_j)$.
If $v_j=p(u)$ the two values $T_j(u)$ and $T_{j+1}(u)$ are equal as the corresponding change in stub length has no effect on the children of $u$.
Computing $\solshort(u)$ and $\sollong(u)$ clearly takes $O(k)$ time.

>>>>

For a vertex $u \in V$ and a neighbor $v$ in $C$ there is a well-defined stub length $\ell$ of $u$

@@

%\begin{proof}
	We reduce from the \NP-hard problem \ppsat~\cite{l-pftu-82} using similar ideas as in Bruckdorfer's general proof of the hardness of \maxsped~\cite{b-sgh-15}. %, which reduces from the more restricted problem \textsc{Planar Positive 1-in-3 Sat}. 
	Let $\phi$ be a planar 3-Sat formula with $n$ variables $\{x_1, \dots, x_n\}$ and $m$ clauses $\{c_1, \dots, c_m\}$, each consisting of three literals.
	We can assume that $\phi$ comes with a planar drawing of its variable-clause graph $H_\phi$ that has a vertex for each variable $x_i$ placed on a horizontal line and a vertex for each clause $c_j$ either above or below the horizontal line and connected on that side of the line to the three variables appearing in $c_j$. 
	In our gadget-based reduction we mimic the drawing of $H_\phi$ by creating a 3-plane drawing $\Gamma_\phi$ as a set of line segments of uniform length and a value $L$ such that $\Gamma_\phi$ has a SPED with ink at least $L$ if and only if $\phi$ is satisfiable.
	Figure~\ref{fig:gadget_connection} shows a sketch of the gadgets.

		\begin{figure}
%	\begin{wrapfigure}[15]{r}{0.55\textwidth}
			\centering
			\includegraphics{Reduction_visual}
			\caption{Example of three variable gadgets and a satisfied clause gadget. Dotted parts do not belong to the SPED.}
			\label{fig:gadget_connection}
%	\end{wrapfigure}		
		\end{figure}
	All segments in the gadgets are of length 5.
	We use pairs of intersecting segments, alternatingly colored red and green.
	The intersection point of each red-green segment pair is at distance 1 from an endpoint.
	Thus the maximum amount of ink contributed by such a pair is 7 (one full segment of length 5 and the other one with two stubs of length 1 each).

@@

	Conversely, assume that we have a SPED with ink value $L$. 
	By construction, every red-green segment pair and every clause gadget must contribute its respective maximum ink value.
	In particular, each variable gadget is either in state \emph{true} or \emph{false}.
	We claim that this truth assignment satisfies $\phi$.
	For each clause, one segment must have full length and as argued before the path connecting that segment to a variable gadget must have all segments at even positions as length-1 stubs and all segments at odd positions as full segments. 
	In particular the first segment of the literal path is fully drawn and therefore in a SPED with ink $L$ the intersecting segment of the variable cannot be fully drawn.
	In fact, it must be drawn as length-1 stubs for otherwise the variable gadget would not contribute its maximum ink value.
	But then by construction the selected variable value sets our literal to \emph{true}.
%\end{proof}

@@

Section~\ref{sec:hardness} showed that \maxsped is generally \NP-hard for $k \ge 3$. 
Now we consider the special case that the intersection graph of the $k$-plane input drawing is a tree or a cactus. In both cases we present polynomial-time dynamic programming algorithms. Let $ C = (V,E) $ be the intersection graph of a given drawing $\Gamma$ of graph $ G $ as defined in Section~\ref{sec:preliminaries}. 
Let $ u \in V $ and $ \deg(u) = \delta $. 
Then for the corresponding segment $ s(u) \in S $ there are $ \delta + 1 $ relevant stub pairs including the whole segment, see Fig.~\ref{fig:tree}. 
Let $ \ell_1(u), \dots, \ell_\delta(u) \in \mathbb{R}_+$ be the stub lengths induced by the intersection points of $s(u)$ with the segments of the neighbors of $u$,  sorted from shorter to longer stubs. We define  $ \ell_0(u) $ as the length of the whole segment $ s(u) $.

\begin{figure}[tbp]
	\centering
 	\includegraphics{tree_illustration}
	\caption{A segment $s(u)$ with five intersecting segments and the different induced stub lengths. The boxed stub lengths are considered in $\solshort(u)$ and do not affect $p(u)$.} % The shaded areas are the stubs, $ p(u) $ is the parent of $ u $. The stubs in the boxed area contribute to $ \solshort(u) $.}
	\label{fig:tree}
\end{figure}

@@

Figure~\ref{fig:tree} highlights the stub lengths that are considered in $\solshort(u)$. We recursively define % for a vertex $ u $ and stub length $ \ell_i(u) $:

@@

Recurrence~(\ref{rec:tree}) can be solved naively in $ O(mk^2) $ time, where $m = |V|$. 
Using the order on the stub lengths  we can improve this to $ O(mk) $ time by computing all $ T_i(u) $ for one $ u \in V $ in $O(k)$ time.
Let $u \in V$ be a vertex with degree $\deg(u) = \delta$. 
The values $T_0(u) = \ell_0(u) + \sum_{v\in c(u)} \solshort(v)$ and $T_1(u) = \ell_1(u) + \sum_{v\in c(u)} \sollong(v)$ for the whole segment $s(u)$ and the shortest stubs can be computed in $O(k)$ time each.
Now $T_{j+1}(u)$ can be computed from $T_j(u)$ in $O(1)$ time as follows.
Let $v_j $ be the neighbor of $u$ that induces stub length $\ell_j(u)$ and assume $v_j \ne p(u)$.
In $T_j(u)$ we could still count the value $\sollong(v_j)$, but in $T_{j+1}(u)$ the stub length of $u$ implies that $v_j$ can contribute only $\solshort(v_j)$. 
Then $T_{j+1}(u) = T_j(u) - \sollong(v_j) + \solshort(v_j)$.
If $v_j=p(u)$ the two values $T_j(u)$ and $T_{j+1}(u)$ are equal as the corresponding change in stub length has no effect on the children of $u$.
Computing $\solshort(u)$ and $\sollong(u)$ clearly takes $O(k)$ time.
